Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-04-03 overall.tex
Fix some more numbers based on the new measure of number of cores

We evaluate the efficiency, scalability and reliability of the deployment of PanDA WMS on Titan by characterizing the behavior of both PanDA Broker and AthenaMP.
% two dimensions: (i) backfill utilization by the PanDA Brokers; and (ii) the
% amount of runtime spent performing detector simulations.
We based our characterizations on measuring the number of cores utilized by
ATLAS on Titan, the overall backfill availability, the number of detector
simulations performed on all the resources available to ATLAS and on Titan, and
the failure rate of these simulations. We also consider the the time taken to
setup AthenaMP and to simulate events in the ATLAS detector with it. All the
measurements were performed between January 2016 and February 2017, hereafter
called `experiment time window'.

\mtnote{TODO: expand the previous paragraph to include dimensions of AthenaMP characterization.}

@@

Figure~\ref{fig:backfill-utilization} (gray bars) shows the number of
core-hours used by ATLAS during the experiment time window. ATLAS consumed a
total of 73.8M core-hours, for an average of $\approx$7M core-hours a month,
with a minimum of 3.3M core-hours in April 2016 and a maximum 14.8M core-hours
in February 2017.

@@

Figure~\ref{fig:backfill-utilization} (green bars) shows backfill utilization
during the experiment time window.  Efficiency
(Figure~\ref{fig:backfill-utilization}, red labels) is defined as the fraction
of Titan core-hours utilized by the PanDA Brokers (for ATLAS jobs) to the total
of Titan’s backfill availability. The brokers reached 18\% average efficiency,
with a minimum 8.9\% efficiency on April 2016 and a maximum 33.5\% efficiency on
February 2017. The total backfill available was 38.1M Titan core-hours in April
2016, and 33.1M Titan core-hours in February 2017. This shows that the measured
difference in efficiency did not depend on a comparable difference in total
backfill availability.

\mtnote{Elaborate further and put it at the end, after giving several of the
measurements we would need. ``Currently, this measure of efficiency is being
refined on the base of the actual minimum amount of walltime that PanDA Broker
can request when submitting job requests. This is not a fixed value as it
depends on variable initialization and configuration overheads, simulation time,
and number of nodes obtained. Part of the undergoing effort is to enable the
measurements of  the distribution of these parameters to derive an optimization
function.''}

\jhanote{I would plot the efficiency on the Y2 axis for both
[diagrams].}\mtnote{Done but then the first diagram ended up plotting a subset
of the data of the second diagram. I used only the second diagram.} \jhanote{I
would suggest similar style of X-axis as previous figure. Difficult to parse
which is April if we are going to call out a specific month. Also consistency in
style is generally good to keep cognitive burden low.}\mtnote{Done.}

@@

During the experiment time window, about 2.25M detector simulation jobs were
completed on Titan, for a total of 225M events processed. This is equivalent to
0.9\% of all the 250M detector simulations performed by ATLAS in the same period
of time, and 3.5\% of the 6.6B events processed by those jobs. Comparatively,
Titan contributed 3.9\% of the total of around 200K cores available to ATLAS on
Grid, Cloud, and HPC infrastructures together. These figures confirms the
relevance of supercomputers' resource contribution to the LHC Run 2, especially
when accounting for the amount of unused backfill availability and the
improvement of PanDA efficiency across the experiment time window.

>>>>

We evaluate the efficiency, scalability and reliability of the deployment of
PanDA WMS on Titan by characterizing the behavior of both PanDA Broker and
AthenaMP. All the measurements were performed between January 2016 and February
2017, hereafter called `experiment time window'.

% two dimensions: (i) backfill utilization by the PanDA Brokers; and (ii) the
% amount of runtime spent performing detector simulations.

% We based our characterizations on measuring the number of cores utilized by
% ATLAS on Titan, the overall backfill availability, the number of detector
% simulations performed on all the resources available to ATLAS and on Titan,
% and the failure rate of these simulations. We also consider the the time taken
% to setup AthenaMP and to simulate events in the ATLAS detector with it. All
% the measurements were performed between January 2016 and February 2017,
% hereafter called `experiment time window'.

% \mtnote{TODO: expand the previous paragraph to include dimensions of AthenaMP
% characterization.}

@@

% Figure~\ref{fig:backfill-utilization} (gray bars) shows the number of
% core-hours used by ATLAS during the experiment time window. ATLAS consumed a
% total of 73.8M core-hours, for an average of $\approx$7M core-hours a month,
% with a minimum of 3.3M core-hours in April 2016 and a maximum 14.8M core-hours
% in February 2017.

Figure~\ref{fig:backfill-utilization} (gray bars) shows the number of core-hours
used by ATLAS during the experiment time window. ATLAS consumed a total of 39.3M
core-hours, for an average of $\approx$3.7M core-hours a month, with a minimum
of 1.7M core-hours in April 2016 and a maximum 7.8M core-hours in February 2017.

@@

Figure~\ref{fig:backfill-utilization} (blue bars) shows backfill utilization
during the experiment time window. Efficiency
(Figure~\ref{fig:backfill-utilization}, red labels) is defined as the fraction
of Titan core-hours utilized by the PanDA Brokers (for ATLAS jobs) to the total
of Titan’s backfill availability. The brokers reached 18\% average efficiency,
with a minimum 8.9\% efficiency on April 2016 and a maximum 33.5\% efficiency on
February 2017.
% The total backfill available was 38.1M Titan core-hours in April 2016, and
% 33.1M Titan core-hours in February 2017. This shows that the measured
% difference in efficiency did not depend on a comparable difference in total
% backfill availability.
The total backfill available was 20.3M Titan core-hours in April
2016, and 17.6M Titan core-hours in February 2017. This shows that the measured
difference in efficiency did not depend on a comparable difference in total
backfill availability.

% \mtnote{Elaborate further and put it at the end, after giving several of the
% measurements we would need. ``Currently, this measure of efficiency is being
% refined on the base of the actual minimum amount of walltime that PanDA Broker
% can request when submitting job requests. This is not a fixed value as it
% depends on variable initialization and configuration overheads, simulation time,
% and number of nodes obtained. Part of the undergoing effort is to enable the
% measurements of  the distribution of these parameters to derive an optimization
% function.''}

% \jhanote{I would plot the efficiency on the Y2 axis for both
% [diagrams].}\mtnote{Done but then the first diagram ended up plotting a subset
% of the data of the second diagram. I used only the second diagram.} \jhanote{I
% would suggest similar style of X-axis as previous figure. Difficult to parse
% which is April if we are going to call out a specific month. Also consistency
% in style is generally good to keep cognitive burden low.}\mtnote{Done.}

@@

During the experiment time window, about 2.25M detector simulation jobs were
completed on Titan, for a total of 225M events processed. This is equivalent to
0.9\% of all the 250M detector simulations performed by ATLAS in the same period
of time, and 3.5\% of the 6.6B events processed by those jobs. Comparatively,
Titan contributed 3.9\% of the total of around 200K\mtnote{TODO: Fix this
number} cores available to ATLAS on Grid, Cloud, and HPC infrastructures
together. These figures confirms the relevance of supercomputers' resource
contribution to the LHC Run 2, especially when accounting for the amount of
unused backfill availability and the improvement of PanDA efficiency across the
experiment time window.
